---
title: "SI618 HW2"
author: "Joshua Gardner"
date: "March 16, 2016"
output: html_document
---

#Loading and Cleaning Data (5 points)
First the TSV data file created in part 1 is loaded into a R data frame using the **read.table()** function. The city, state and main\_category columns should be converted to factors. Then listwise deletion (http://en.wikipedia.org/wiki/Listwise_deletion) is applied to remove records with missing data (use the **na.omit()** function). Here is the summary of the data frame:
```{r, echo=FALSE}
yelpdata = read.table("si618_hw2_jpgard_output.tsv", sep = '\t', header = TRUE, quote = "", comment.char = "", stringsAsFactors = FALSE)
yelpdata = na.omit(yelpdata)
yelpdata$city = factor(yelpdata$city)
yelpdata$state = factor(yelpdata$state)
yelpdata$main_category = factor(yelpdata$main_category)
summary(yelpdata)
```


#Histograms of Star Ratings (10 points)
Histograms of star ratings are plotted with the **qplot()** function. Both actual counts and density plot are shown. (Use **binwidth=1)**

```{r, echo=FALSE, fig.width=12.5, fig.height=5}
library(ggplot2)
ggplot(yelpdata, aes(x = stars, fill = state)) + geom_histogram(binwidth = 1, alpha = 0.5) + facet_grid(. ~ state) + scale_x_continuous(breaks = c(1,2,3,4,5)) + scale_alpha(guide = 'none') + xlab("Stars")
#qplot(stars, data = yelpdata, facets = .~state, geom = "histogram", binwidth = 1, fill = state, alpha = 0.5, main = "Histograms of Star Ratings") + xlim(1,6) + scale_alpha(guide = 'none') + scale_x_continuous(breaks = c(1,2,3,4,5))
ggplot(yelpdata, aes(x = stars, fill = state)) + geom_histogram(aes(y = ..density..), binwidth = 1, alpha = 0.5) + facet_grid(. ~ state) + xlim(1, 6) + scale_x_continuous(breaks = c(1,2,3,4,5)) + ggtitle("Histograms of Star Ratings") + scale_alpha(guide = 'none') + xlab("Stars")
```

#Histograms of Review Counts (10 points)
Histograms of review counts are plotted with the **qplot()** function. (Use **binwidth=10**)

```{r, echo=FALSE, fig.width=9, fig.height=6}
ggplot(yelpdata, aes(x = review_count)) + geom_histogram(binwidth = 10, alpha = 0.5) + xlab("Review Counts") + ggtitle("Histograms of Review Counts") + scale_alpha(guide = 'none')


```

We can see that the distribution of review counts has a long tail. To zoom in on the bars to the left of the 200 mark, we use the **subset()** function to select just the data with review count <= 200. And then plot the histogram again with **binwidth=1**.

```{r, echo=FALSE, fig.width=9, fig.height=6}
yelpdata2 = subset(yelpdata, review_count <= 200)
ggplot(yelpdata2, aes(x = review_count)) + geom_histogram(binwidth = 1, alpha = 0.5) + xlab("Review Counts") + ggtitle("Histograms of Review Counts (Filtered)") + scale_alpha(guide = 'none')
```

#Boxplot of Star Ratings by States (10 points)

```{r, echo=FALSE, fig.width=9, fig.height=6}
#ggplot(yelpdata, aes(x = state)) + geom_boxplot(aes(y = stars, colour = state, fill = state, alpha = 0.5))

qplot(state, stars, data = yelpdata, geom = "boxplot", colour = state, fill = state, alpha = 0.5, main = "Star Ratings By States") + scale_alpha(guide = 'none') + ylab("Stars") + theme(axis.title.x=element_blank())
```

#Jittered Plot of Star Ratings by States (10 points)

```{r, echo=FALSE, fig.width=9, fig.height=6}
#ggplot(yelpdata, aes(x = state)) + geom_boxplot(aes(y = stars, colour = state, fill = state, alpha = 0.5))

qplot(state, stars, data = yelpdata, geom = "jitter", colour = state, fill = state, alpha = 0.5, main = "Star Ratings By States") + scale_alpha(guide = 'none') + ylab("Stars") + theme(axis.title.x=element_blank())
```

#Bar Chart of Number of Businesses by State (10 points)

The states should be orderd by descreading height of bars.

```{r, echo=FALSE, fig.width=9, fig.height=6}

qplot(reorder(state, state, function(x)-length(x)), data = yelpdata, alpha = 0.5) + scale_alpha(guide = 'none')
```

#Jittered Scatterplot of Stars and Review Counts (10 points)

```{r, echo=FALSE, fig.width=9, fig.height=6}
qplot(stars, review_count, data = yelpdata, colour = state, geom = "jitter", alpha = 0.5, size = 1) + scale_alpha(guide = 'none') + scale_size(guide = 'none')
```


#Slice and Dice Data with plyr

##Subsetting Data (10 points)

We first rank the business in each city for each main category. Then top 5 ranked businesses in each city for each main category are found. We are interested in is the top 5 business with main category of “Chinese” in each city. The data should be ordered by city names, and then by ranks. The result is listed below.

```{r, echo=FALSE}
library(plyr)
chinese = subset(yelpdata, main_category == "Chinese")
chineseranks = ddply(chinese, .(city), transform, rank = rank(-stars, tie = "first"))
chinesetop5 = subset(chineseranks[,c(2,1,7,4)], rank <= 5)
arrange(chinesetop5, city, rank)

#in single line:
#arrange(subset(ddply(chinese, .(city), transform, rank = rank(-stars, tie = "first"))[,c(2,1,7)], rank <= 5), city, rank)
```

#Summarize Data (10 points)
Next, we compute the mean review counts of all businesses for each state and plot the bar chart below.

```{r}
meanreviews = ddply(yelpdata, .(state), summarise, mean_reviews = mean(review_count))
qplot(reorder(state, mean_reviews, function(x)-x), mean_reviews, data = meanreviews, geom = "bar", stat = "identity", alpha = 0.5) + scale_alpha(guide = 'none') + ylab("Mean Review Counts")
```








